---
title: "Introduction to Double DiD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{double}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we show how to deal with confounding events, see [Tsai (2024)](https://arxiv.org/abs/2409.05184) for the full treatment. We first simulate some data with confounding events. 

```{r setup}
library(fastdid)
library(ggplot2)
library(data.table)
simdt <- sim_did(1e+04, 5, cov = "cont", hetero = "all", balanced = TRUE, seed = 1, 
                 second_cohort = TRUE, second_het = "no") #comfounding event
dt <- simdt$dt #dataset

#ground truth att
att <- simdt$att |> merge(dt[,.(w = .N),by = "G"], by = "G")
att[, event_time := time-G]
att <- att[event == 1,.(att = weighted.mean(attgt, w)), by = "event_time"]
```

Using the default estimator, the estimates (black line) is biased from the ground truth (red line).

```{r, naive}
naive_result <-fastdid(data = dt, 
                 timevar = "time", cohortvar = "G", unitvar = "unit", 
                 outcomevar = "y", result_type = "dynamic")
plot_did_dynamics(naive_result) + 
  geom_line(aes(y = att, x = event_time), 
            data = att, color = "red") + theme_bw()
```
Diagnostics can be obtained by using `time >= G2` as outcome. As we set the effect of the confounding event at 10 constantly, we can see that the bias of the default estimator is roughly 10 times the diagnostics.

```{r, diag}
dt[, D2 := time >= G2]
diag <- fastdid(data = dt, 
                 timevar = "time", cohortvar = "G", unitvar = "unit",
                 outcomevar = "D2", result_type = "dynamic")
plot_did_dynamics(diag) + theme_bw()
```

Using the double did estimator with `cohortvar2`, the estimator recovers the ground truth.

```{r, double}
double_result <-fastdid(data = dt, 
                 timevar = "time", cohortvar = "G", unitvar = "unit",
                 outcomevar = "y", result_type = "dynamic",
                 cohortvar2 = "G2", event_specific = TRUE)
plot_did_dynamics(double_result) + 
  geom_line(aes(y = att, x = event_time), 
            data = att, color = "red") + theme_bw()
```

Double DiD also allow for two additional aggregation scheme: group-group-time ("group_group_time") and dynamic-staggered ("dynamic_stagger", event time by event stagger, G1-G2).

```{r, agg}
double_result_ds <-fastdid(data = dt, 
                 timevar = "time", cohortvar = "G", unitvar = "unit",
                 outcomevar = "y", result_type = "dynamic_stagger",
                 cohortvar2 = "G2", event_specific = TRUE)

double_result_ggt <-fastdid(data = dt, 
                 timevar = "time", cohortvar = "G", unitvar = "unit",
                 outcomevar = "y", result_type = "group_group_time",
                 cohortvar2 = "G2", event_specific = TRUE)

```

## Effect models

The second stage of double DiD imputes the effect of the first event into the periods after the confounding event. The default imputation is the parallel treatment effects assumption. `effect_model` replaces it with a formula for the effect. The formula is fit on the clean cells, the group-times before the confounding event, and imputed into the confounded cells. A cell is reported only when its regressor row is in the row space of the fit, so the output never extrapolates through a rank-deficient fit.

The parallel treatment effects assumption is the two-way model below, fit on every clean cell at once. A quadratic profile in event time with a cohort level is a stronger model that extrapolates within a cohort. `"unrestricted"` reports the clean cells only.

```{r, effect}
parallel_fit <- fastdid(data = dt,
                 timevar = "time", cohortvar = "G", unitvar = "unit",
                 outcomevar = "y", result_type = "dynamic", cohortvar2 = "G2",
                 effect_model = ~ factor(gvec) + factor(gactive):factor(t))

quadratic_fit <- fastdid(data = dt,
                 timevar = "time", cohortvar = "G", unitvar = "unit",
                 outcomevar = "y", result_type = "dynamic", cohortvar2 = "G2",
                 effect_model = ~ factor(gvec) + poly(e, 2))

clean_only <- fastdid(data = dt,
                 timevar = "time", cohortvar = "G", unitvar = "unit",
                 outcomevar = "y", result_type = "dynamic", cohortvar2 = "G2",
                 effect_model = "unrestricted")
```

With `full = TRUE`, `effect_diag` holds the weight of every first-stage cell in every reported cell, the estimability and the extrapolation gap of every candidate cell, the residuals of the fit, and an over-identification test of the model on the clean cells.

```{r, diag2}
full_fit <- fastdid(data = dt,
                 timevar = "time", cohortvar = "G", unitvar = "unit",
                 outcomevar = "y", result_type = "group_group_time", cohortvar2 = "G2",
                 effect_model = ~ factor(gvec) + poly(e, 2), full = TRUE)
head(full_fit$effect_diag$y$cells)
full_fit$effect_diag$y$overid$target
```

When the events are the first and the second occurrence of one event kind, set `effect_fit = "ordered"`. The cohort dates must increase. The design row of a cell is then the sum over its active events, so a formula can share the event-time profile across occurrences. `result_type = "dynamic_event"` averages the effect of every occurrence at each event time.

```{r, ordered, eval = FALSE}
ordered_fit <- fastdid(data = dt_ordered,
                 timevar = "time", cohortvar = "G", unitvar = "unit",
                 outcomevar = "y", result_type = "dynamic_event", cohortvar2 = "G2",
                 effect_model = ~ 0 + factor(gvec):factor(event) + factor(e),
                 effect_fit = "ordered")
```
